Analyzing Cooperation with Game Theory and Simulation
========================================================================
Task Set 1: Noisy Prisoner Dilemma
========================================================================

**Date: 2013-10-11**
**Author: Sebastian Kranz (sebastian.kranz@uni-ulm.de)**

### Task 1.1 Install software and work through the tutorial

First read the tutorial for the seminar and for the R package StratTourn, install the software and try out the examples in the tutorial.

### Task 1.2 Develop strategies for 4 scenarios of a Noisy Prisoners' Dilemma Game

In the PD game of the tutorial, players can perfectly observe the action that their partner has chosen in the previous period. This is a strong assumption. In many real world problems that share some structure with a PD game, it is not always easy to observe what the other player has done (think of environmental agreements to reduce polution or reduce quantity of fishing, or hard to observe work effort on joint projects) Resulting information problems may become a severe obstacle to efficient cooperation and strategies have to take them into account.

#### Noisy PD Game

In this task we consider noisy prisoners' dilemma games that add a simple form of information problem. There is an exogenously given probability *err.D.prob* that players observe that a player has chosen D in the previous period irrelevant of his actual choice (which might as well be C). Similarly, there can be a positive probability *err.C.prob* that a player's action is observed as C, even though he has chosen D. Formally, the period t+1 observation $obs_{i,t+1}$ of player i's period t action $a_{i,t}$ is distributed as follows:
\[
obs_{i,t+1}=
\begin{cases}
  a_{i,t} & \mbox{with prob 1-err.D.prob-err.C.prob}\\
  D & \mbox{with prob err.D.prob}\\
  C & \mbox{with prob err.C.prob}
\end{cases}
\]

Both players always make the same observations, i.e. each player knows what the other player has observed. That all players make the same noisy observations is called *imperfect public monitoring* in game theoretic language. (If there are no observation errors, we have *perfect monitoring*. If players could make different observations, we would have *private monitoring*)

Every team shall develop strategies for each of the following 4 scenarios (one strategy per scenario):

#### 1. Baseline scenario:

  * err.D.prob = 0.15
  * err.C.prob = 0
  * $\delta=0.95$
  
#### 2. High err.D.prob scenario:

  * err.D.prob = 0.3
  * err.C.prob = 0
  * $\delta=0.95$

#### 3. err.C.prob scenario:

  * err.D.prob = 0.15
  * err.C.prob = 0.3
  * $\delta=0.95$

#### 4. Low $\delta$ scenario:

  * err.D.prob = 0.15
  * err.C.prob = 0
  * $\delta=0.5$


For each scenario separately the strategies of all teams will compete in a tournament as described in Section 4 of the tutorial. Scores are computed with the formula
\[
    score = efficiency - 2*instability - 20*instability^2
\]


#### Two stages of the tournament  
As described in Section 4 of the tutorial, the tournament will have a second stage. After we run the first stage, every team gets the source code of all strategies and has 1-2 weeks time to develop strategies that have the only goal to increase the instability of competing strategies. The final score of the initially submitted strategies will be computed after this second stage.

#### A code example for testing strategies
```{r include=FALSE}
library(StratTourn)
set.storing(FALSE)
runif(1)
```

```{r cache=!TRUE}
library(StratTourn)

# scenario 1: baseline
err.D.prob = 0.15; err.C.prob = 0; delta = 0.95
# scenario 2: high err.D.prob
err.D.prob = 0.3; err.C.prob = 0; delta = 0.95
# scenario 3: err.C.prob 
err.D.prob = 0.15; err.C.prob = 0.3; delta = 0.95
# scenario 4: low delta
err.D.prob = 0.15; err.C.prob = 0; delta = 0.5


# Current scenario
# baseline scenario
err.D.prob = 0.15; err.C.prob = 0; delta = 0.95


game = make.pd.game(err.D.prob=err.D.prob, err.C.prob=err.C.prob)

# Pick a pair of strategies 
strat = nlist(tit.for.tat,grim.trigger)

# Let the strategies play against each other
run.rep.game(delta=delta, game=game, strat = strat)

# Init and run a tournament of several strategies against each other  
strat = nlist(grim.trigger,tit.for.tat,always.defect, always.coop)
tourn = init.tournament(game=game, strat=strat, delta=delta,
                        score.fun = "efficiency - 2*instability -20*instability^2")
tourn = run.tournament(tourn=tourn, R = 10)
tourn

```
