---
title: "Ranking of strategies"
tab_label: "matches" 
output: html_document
---

```{r "init_data", include=FALSE}
library(StratTourn)
library(shiny)
library(xtable)
library(knitr)
library(ggvis)
library(dplyr)
setwd("D:/libraries/StratTourn")

tourn = load.tournament("Tourn_Noisy_PD_20140721_202445.Rdata")
# Data for each match
md = tourn$dt
md = add.other.var(md,c("strat","u"))
md$delta.u = md$u - md$other.u
# Names of all strategies
strats = unique(md$strat)

used.strats = strats
shown.strats = strats

```

```{r "ggvis_duels", echo=FALSE, results='asis', fig.width=10, output.type="ggvis"}
d= get.matches.vs.grid(md)
if (TRUE) {
  d$date=2001
  state = '{"orderedByY":false,"showTrails":true,"xZoomedDataMin":-1,"colorOption":"4","xAxisOption":"2","yAxisOption":"3","sizeOption":"_UNISIZE","xZoomedIn":false,"xLambda":1,"iconKeySettings":[],"yZoomedIn":false,"playDuration":15000,"xZoomedDataMax":2,"orderedByX":false,"iconType":"BUBBLE","time":"2001","uniColorForNonSelected":false,"yZoomedDataMin":-1,"yZoomedDataMax":2,"dimensions":{"iconDimensions":["dim0"]},"nonSelectedAlpha":0.4,"duration":{"timeUnit":"Y","multiplier":1},"yLambda":1}'
  
  p = gvisMotionChart(d, idvar = "pair", timevar = "date", xvar = "u1",
  yvar = "u2", colorvar = "strat1", sizevar = "share1x2", 
  options = list(state=state))

  #plot(p)
  print(p, tag="chart")
  
}
if (!TRUE) {
library(ggplot2)
qplot(data=d, x=u1,y=u2, color=strat1,shape=strat2, geom="point",size=I(4), main="Payoffs of strategy duels")
}


if (!TRUE) {
library(ggvis)

tooltip_fun <- function(x) {
  if(is.null(x)) return(NULL)
  paste0(x$strat1, " vs ", x$strat2,"\n<br>\n",round(x$u1,2), " vs ",round(x$u2,2))
}

gg = ggvis(data=d, x=~u1, y=~u2,stroke=~strat1,fill=~strat1, shape=~strat2) %>%
  layer_points() %>% 
  add_legend(c("fill","stroke", "shape"), title="strat1: color vs strat2: shape") %>%
  add_tooltip(tooltip_fun, "hover")

#print(gg)
gg
}

```

The figure illustrates the payoffs of all strategy duels. The color represent strategy 1, by hovering over a bubble you also see the strategy 2.

Hints for interpreting the graph:

  - We can compare different corner areas in which a bubble lies:
    - top-right corner: effective cooperation, both strategies have high payoffs
    - bottom-left corner: mutual destruction, both strategies get low payoffs
    - top-left corner: strategy 1 is exploited by strategy 2  
    - bottom-right corner: strategy 1 exploits strategy 2
    
  - A line of bubbles of the same color allows to compare a strategy's outcomes against all other strategies. The shape of the line can be interpreted:
    - A positive slope of strategy's bubbles suggest that the strategies behaves reciprocal: a higher own payoff corresponds to a higher payoff of other strategies.
    - Negative slope suggests that the strategy is either exploited or exploits some strategies
  
  - In a symmetric game, the outcomes when a strategy plays against itself should lie roughly on a 45° degree line. Small variations are possible due to random factors that influence in the outcomes for player 1 and 2.
  
  - All bubbles lie in the convex hull of the stage game payoffs 


